---
title: "p8105_hw2_to2345"
author: Tianna_Ouyang
date: 2019/09/24
output: html_document
---

```{r setup,message=FALSE}
# Load the packages that will be used in this assignment
knitr::opts_chunk$set(echo = TRUE)
library(psych)
library(tidyr)
library(tidyverse)
library(readxl)
library(dplyr)
library(knitr)
library(kableExtra)

```

## Problem 1

```{r question1, warning=FALSE}
# Read and clean the Mr. Trash Wheel sheet
trash_data = 
  readxl::read_excel(
    "../data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
    sheet = 1,
    range = "A2:N211"
    ) %>%
  janitor::clean_names() %>%
  na.omit() %>% 
  mutate(
    sports_balls = as.integer(round(sports_balls, 0))
  ) %>% 
  view

# Read and clean precipitation data for 2017 and 2018
precipitation_2018 = 
  readxl::read_excel(
  "../data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
  sheet = "2018 Precipitation",
  range = "A2:B14"
  ) %>% 
  mutate(year = '2018') %>% 
  view

precipitation_2017 = 
  readxl::read_excel(
  "../data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
  sheet = "2017 Precipitation",
  range = "A2:B14"
  ) %>% 
  mutate(year = '2017') %>% 
  view

precipitation =
  bind_rows(precipitation_2017, precipitation_2018) %>% 
  janitor::clean_names() %>% 
  mutate(
    month = month.name[month]
  ) %>% 
  view
```


* The cleaned Mr.Trash Wheel dataset has `r ncol(trash_data)` variables, `r nrow(trash_data)` records.
* The precipitation data for 2017 and 2018 dataset has `r ncol(precipitation)` variables, `r nrow(precipitation)` records.
* Some basic descriptive datas for the weight of trash in tons from Mr.Trash Wheel Dataset are:
  + Mean: `r mean(pull(trash_data, weight_tons))`
  + Median: `r median(pull(trash_data, weight_tons))`
  + Min: `r min(pull(trash_data, weight_tons))`
  + Max: `r max(pull(trash_data, weight_tons))`
  + Variance: `r var(pull(trash_data, weight_tons))`
  + Skewness: `r skew(pull(trash_data, weight_tons))`

* For available data, the total precipitation in 2018 is `r sum(pull(precipitation, total)[pull(precipitation, year) == 2018], na.rm = TRUE)`;
* The median number of sports balls in a dumpster in 2017 is `r median(pull(trash_data,sports_balls)[pull(trash_data, year) == 2017])`.

## Problem 2

```{r, warning=FALSE}
# import and clean the data in pols-month.csv
pols_month = 
  read_csv("../data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c('year','month','day', sep = '-')) %>% 
  mutate(
    month = month.name[as.integer(month)],
    year = as.integer(year),
    president = recode(prez_gop, '1' = 'gop', '0' = 'dem')) %>% 
  select(-c('prez_dem','prez_gop','-','day')) %>% 
  view
```

```{r, warning=FALSE}
# import and  clean the data in snp.csv
snp = 
  read_csv('../data/fivethirtyeight_datasets/snp.csv') %>% 
  janitor::clean_names() %>% 
  separate(date, into = c('month','day','year', sep = '/')) %>% 
  mutate(
    month = month.name[as.integer(month)],
    year = as.integer(year)
    ) %>%
  select(year, month, close) %>%
  arrange(year, match(month, month.name)) %>% 
  view
```



```{r, warning=FALSE}
# define a function to capitalize the first character in a string (will eb s)
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x}

# import and tidy the unemployment data so that it can be merged with the previous datasets
unemployment = 
  read_csv('../data/fivethirtyeight_datasets/unemployment.csv') %>% 
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = 'month',
    values_to = 'unemployment') %>%
  mutate(
    month = firstup(month),
    month = month.name[match(month, month.abb)]
  ) %>% 
  view

```


*Join the datasets by merging snp into pols, and merging unemployment into the result.*
```{r}
# Join the dataset imported with common keys into new dataset 
result = left_join(pols_month, snp, by = c('year' = 'year','month' = 'month'))
result = left_join(result, unemployment, by = c('year' = 'year', 'month' = 'month'))
```


*Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).*

* The pols-month dataset has `r ncol(pols_month)` variables, `r nrow(pols_month)` records. 
  + The variables in this dataset are `r names(pols_month)`.
* The snp dataset has `r ncol(snp)` variables, `r nrow(snp)` records.
  + The variables in the dataset are `r names(snp)`.
* The unemployment dataset has `r ncol(unemployment)` variables, `r nrow(unemployment)` records.
  + The variables in the dataset are `r names(unemployment)`
* The resulting dataset has `r ncol(result)` variables, `r nrow(result)` records.
  + The range of years in the resulting dataset is [`r range(pull(result, year))`]
  + The variables in the resulting dataset are: `r names(result)`.


## Problem 3

*the names of a categorical predictor and the case structure of string variables changed over time; you’ll need to address this in your data cleaning.*
*Also, some rows seem duplicated, and these will need to be removed.*
```{r}
# import the popular baby names file and perform corresponding data cleaning
baby_name = 
  read_csv("../data/Popular_Baby_Names.csv", col_types = "icccii") %>% 
  janitor::clean_names() %>%
  mutate(
    gender = str_to_lower(gender),
    ethnicity = str_to_lower(ethnicity),
    childs_first_name = str_to_lower(childs_first_name)
  ) %>%
  distinct() %>% 
  view
```

*Produce a table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year.*

```{r}
# make a new table based on the data subsetting from original dataframe with condition
olivia = subset(baby_name, childs_first_name == "olivia" & gender == "female")
olivia = select(olivia, year_of_birth, ethnicity, rank)
kable(olivia, caption = "The popularity of 'Olivia' in female baby over time") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```

*for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis)*

```{r}
# make a scatter plot describing the relationship between rank and the count of names
scatter_point = subset(baby_name, gender == "male" & ethnicity == "white non hispanic" & year_of_birth == 2016)
plot_df = tibble(
  x = pull(scatter_point, rank),
  y = pull(scatter_point, count)
)
ggplot(plot_df, aes(x = x, y = y))+ geom_point(color = "#00AFBB", size = 2, shape = 23)
```

*  From the scatter plot we can see that there is a negative relationship between the number of childen with a name and the rank in popularity of that name, in other words, the higher the number of childer with a name, the rank in popularity of that name will be smaller for male, white non-hispanic children in 2016.

